### 2024-02-13 outputs/paper/paper.Rmd
update

---
title: "The Impact of Past Behavior Normality on Regret: A Replication Study"
subtitle: "Reproduction of 'The Puzzle of Falling US Birth Rates since the Great Recession' (Kearney, Levine & Pardue, 2022)"
author: Heng Ma & Bolin Shen
thanks: "Code and data are available at: https://github.com/Brian031205/paper_2.git. A replication of various aspects in this paper are available at: ______________    "
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Norm theory (Kahneman & Miller 1986) described a tendency for people to associate stronger regret with a negative outcome when it is a result of an exception (abnormal behavior) compared to when it is a result of routine (normal behavior). In this paper, we conducted a replication of one classic experiments on past behavior exception/routine contrasts. We successfully replicated Kahneman and Miller’s (1986) experiments with the classic hitchhiker-scenario (Part 1) and car accident-scenario (Part 2). In both cases, participants examined negative outcomes and tended to indicate a protagonist who deviated from own past behavior as more regretful than another who followed routine. In robbery scenario experiments using compensatory measures (Part 3), there is no significant difference between special and regular situations in the compensation comparison of robbery victims. However, it is showed that robberies in exceptional circumstances are considered more regrettable than robberies in regular circumstances. We discuss implications for current and future research."
output:
  bookdown::pdf_document2
toc: TRUE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.pos = "h", out.extra = "")

# Load libraries
if(!require(Rcpp)){install.packages('Rcpp', dependencies = TRUE)}
#if(!require(jmv)){install.packages('jmv', repos=c('https://repo.jamovi.org', 'https://cran.r-project.org'))}
if(!require(jmv)){install.packages('jmv', dependencies = TRUE)}
if(!require(rstudioapi)){install.packages('rstudioapi', dependencies = TRUE)}
if(!require(effsize)){install.packages('effsize', dependencies = TRUE)}
if(!require(Hmisc)){install.packages('effsize', dependencies = TRUE)}
if(!require(psych)){install.packages('psych', dependencies = TRUE)}
if(!require(ggplot2)){install.packages('ggplot2', dependencies = TRUE)}
if(!require(dplyr)){install.packages('dplyr', dependencies = TRUE)}
if(!require(MBESS)){install.packages('MBESS', dependencies = TRUE)}

library(MBESS)
library(psych)
library(ggplot2)
library(dplyr) 
library(Hmisc)
library(effsize)
library(rstudioapi)
library(jmv)
library(Rcpp)

# Read in data

```

# Introduction

Regret is a pervasive and intense negative emotion, plays a pivotal role in shaping individuals' daily experiences. The recognition that other decisions or circumstances could have led to more favorable outcomes triggers regret (Byrne, 2016; Roese & Epstude, 2017; Kahneman & Tversky, 1982). Influential work by Kahneman and Miller (1986) in norm theory has highlighted the connection between abnormal behavior, heightened counterfactual thinking, and intensified feelings of regret. Their discussion of the concept of normality included experimental research illustrating various factors influencing the perception and interpretation of normality (Feldman & Albarracín, 2017; Feldman, 2018). This approach often employed past behavior as a reference point to assess the normality of an individual's actions.

Illustrating the effect of past behavior normality, Kahneman and Miller (1986) presented a thought experiment involving a car accident. Mr. White, accustomed to driving home along the same route daily, deviated from his routine one day and found himself in an accident. 82% of participants believed that Mr. White would experience more regret than those involved in similar accidents on his usual route. This bias referenced as the "exceptionality effect," serves as a basis for this study, emphasizing the regularity of past behavior.

Kahneman and Miller's (1986) norm theory provides a psychological classification of events as normal or abnormal, assuming that abnormal events elicit stronger reactions because of the increased accessibility of normal events in memory. This high availability of psychologically simulated alternatives intensifies regret over negative outcomes. Everyday behaviors, more easily recalled and psychologically emphasized, enable individuals to vividly imagine counterfactual scenarios in exceptional circumstances, thus amplifying the regret associated with deviations from past behavior (Kahneman & Miller 1986).

This study aims to directly replicate three experiments that demonstrated the exceptionality effect, representing early investigations into the influence of past behavior normality on regret. Specifically, we investigate hitchhiking and car accident scenarios from Kahneman and Miller (1986) in the first two part and replicate Miller and McFarland's (1986) Experiment 1, featuring a robbery scenario, in the part three to test norm theory. Although there is no documented direct replication of these experiments, the existing citation impact of Kahneman and Miller's work emphasizes the significance of a reexamination and detailed explanation of these effects. The study's findings are expected to contribute to a better understanding of the general nature of the exceptionality effect and the intricate influence of past behavior normality on regret.

we use R [@citeR] for all data wrangling and analysis and R packages tidyverse [@tidy], ggthemes [@ggthemes], ggprism [@ggprism] and patchwork [@patchwork] to produce the figures, kableExtra [@kableExtra] to produce the tables.

# Data
The data we used in this paper are obtained from the paper titled “The impact of past behavior normality on regret: replication and extension of three experiments of the exceptionality effect” (Kutscher & Feldman 2022). We first cleaned the data, and then analyzed them by the open-source statistical programming language R (R Core Team 2022). We used functionalities from tidyverse (Wickham et al. 2019), ggplot2 (Wickham 2016), dplyr (Wickham et al. 2022), readr (Wickham, Hester, and Bryan 2022), tibble (Müller and Wickham 2022), janitor (Firke 2021) and knitr (Xie 2014).

## Proportion for perceived regret, injunctive social norms, descriptive social norms, and negative affect
The first figure presents an overview of the perceived regret, injunctive social norms, descriptive social norms, and negative affect proportions associated with the behaviors of Mr. Smith (routine) and Mr. Jones (exception). In terms of regret, the figure highlights an obvious contrast, with Mr. Smith at 7.9% and Mr. Jones at a much higher 92.1%. A similar pattern emerges for injunctive social norms, where Mr. Smith scores 4.7%, contrasting sharply with Mr. Jones at 95.3%. Descriptive social norms show an obvious preference for routine behavior, as reflected in Mr. Smith's 90.6%, while Mr. Jones lags behind at 9.4%. Negative affect follows a parallel trend, with Mr. Smith at 7.3% and Mr. Jones at 92.7%. This figure vividly illustrates the remarkable differences in perceptions between routine and exceptional behaviors, emphasizing the notable impact on regret and social norm evaluations.

## Proportion for perceived regret and luck
Figure 2, titled "Proportions for Perceived Regret and Luck," provides a clear visual representation of the contrasts in perceived regret and luck between the behaviors of Mr. Adams (routine) and Mr. White (exception). The data shows a large difference in regret percentages, with Mr. Adams scoring at 19%, while Mr. White gets a significantly higher 81%. The figure also reveals the aspect of luck, where Mr. Adams has a proportion of 33.3%, distinctly lower than Mr. White, who gets a larger share at 66.7%. This figure effectively demonstrates the difference in perceptions regarding regret and luck associated with routine and exceptional behaviors, revealing the dynamic relationship between these two factors in the participants' evaluations of the scenarios described.

## Victim compensation and regret (with exception conditions combined)
The last figure titled "Victim Compensation and Regret (with Exception Conditions Combined)," presents an overview of two experiments, each with routine and exception conditions. The top two plots correspond to Experiment 1, while the bottom two plots represent Experiment 2, with error bars indicating standard errors. The scales range from 1 to 5 for compensation and 1 to 10 for regret. In Experiment 1, the routine condition includes 114 participants, with standard deviations of 2.37 for compensation and 1.12 for regret. The self-produced exception condition involves 115 participants, with standard deviations of 2.46 for compensation and 0.80 for regret. For the other-produced exception condition (113 participants), standard deviations are 2.76 for compensation and 0.94 for regret. Combining all exception conditions (228 participants) gets standard deviations of 2.61 for compensation and 0.87 for regret. In Experiment 2, routine conditions involve 116 participants, with standard deviations of 3.40 for compensation and 1.19 for regret. The self-produced exception condition (110 participants) shows standard deviations of 3.50 for compensation and 0.67 for regret. The other-produced exception condition (116 participants) has standard deviations of 3.66 for compensation and 0.98 for regret. Combining all exception conditions (226 participants) results in standard deviations of 3.57 for compensation and 0.86 for regret. These detailed data provide an understanding of victim compensation and regret across various experimental conditions.

```{r birthdata, echo=FALSE}
# Makes table of Ten rows of the birth rate data for years 1980-2020
birth_data %>%
  select(year, brate_all, brate_whitenh, brate_blacknh, brate_hisp, brate_unmarried, brate_married) %>%
  filter(year > 1989) %>%
  slice(1:10) %>%
  kable(
    caption = "Ten rows of the birth rate data for years 1980-2020",
    col.names = c("Year", "All races", "White, non-Hispanic", "Black, non-Hispanic", "Hispanic", "Unmarried", "Married"),
    digits = 1,
    booktabs = TRUE, 
    linesep = ""
  )
```

For the aggregated cohort birth data, the variables are mother's age, birth cohort, number of births and cumulative births. The population data provided by @og gives the population count for women by age, year and state. We group the population count by cohort and mother's age and join this to the cohort birth data, with the first 10 rows of this data shown in table \@ref(tab:birthcohort). We also added a column with the birth years for each cohort for easier interpretation.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Process population data to aggregate by cohort
fig5_pop_data1 <- pop_cohort_data %>%
  select(1:32) %>%
  select(-c(stname)) %>%
  group_by(year) %>%
  summarise(across(everything(),sum)) # https://stackoverflow.com/questions/1660124/how-to-sum-a-variable-by-group

fig5_pop_data1 <- fig5_pop_data1 %>%
  pivot_longer(-c(year), names_to = "mage", values_to = "pop")

fig5_pop_data1 <- fig5_pop_data1 %>%
  mutate(mage = as.numeric(str_replace(mage, "fem", ""))) %>%
  mutate(cohort = year - mage) %>%
  mutate(cohort2 = case_when(cohort >= 1968 & cohort <=1972 ~ 1,
                             cohort >= 1973 & cohort <=1977 ~ 2,
                             cohort >= 1978 & cohort <=1982 ~ 3,
                             cohort >= 1983 & cohort <=1987 ~ 4,
                             cohort >= 1988 & cohort <=1992 ~ 5,
                             cohort >= 1993 & cohort <=1997 ~ 6))

fig5_pop_data1 <- na.omit(fig5_pop_data1)
fig5_pop_data1 <- fig5_pop_data1 %>%
  group_by(cohort2, mage) %>%
  summarise(pop = sum(pop))

# Join dataframes for population and birth data
birth_cohort_data <- full_join(birth_cohort_data, fig5_pop_data1, by=c("mage", "cohort2")) %>%
  mutate(cohort_years = case_when(cohort2 == 1 ~ "1968-1972",
                                  cohort2 == 2 ~ "1973-1977",
                                  cohort2 == 3 ~ "1978-1982",
                                  cohort2 == 4 ~ "1983-1987",
                                  cohort2 == 5 ~ "1988-1992",
                                  cohort2 == 6 ~ "1993-1997"))

```

```{r birthcohort, echo=FALSE}
# Make table of First ten rows of the birth cohort data for 6 birth cohorts
birth_cohort_data %>%
  select(2,6,1,3,4,5) %>%
  slice(1:10) %>%
  kable(
    caption = "First ten rows of the birth cohort data for 6 birth cohorts",
    col.names = c("Cohort #", "Birth Years", "Mother's age", "Number of births", "Cumulative births", "Population Count"),
    digits = 1,
    booktabs = TRUE, 
    linesep = ""
  )
```

# Results

## Overall evaluation in regre depending hitchhiker scenario

```{r table1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Rates for Regret, Social norms and Negative affect"}

Orginal_table <- read.csv("/cloud/project/inputs/data/Ex1.csv", header = TRUE, stringsAsFactors = FALSE, fileEncoding = "UTF-8-BOM")
print(Orginal_table)
```

Figure \@ref(fig:fig1) is a replication of Figure 1 in the original paper, In this section, participants evaluated a classic scenario known as the hitchhiker scenario, to assess the impact of past behavior on the perception of regret following negative outcomes. 92% of the 315 participants believed that Mr. Jones, who almost never gives rides to others, felt a stronger sense of regret. The majority of participants believed that Mr. Jones, who almost never picks up hitchhikers, exhibits behavior that is more common in society (95% out of 326 respondents), whereas Mr. Smith, who frequently gives rides to hitchhikers, is more likely to be criticized by society (91% out of 310 respondents). For the assessment of negative affect, 93% of people believed that Mr. Jones experienced more negative affect compared to Mr. Smith. Overall, these two experiments provided very close replications, similar to the the original study.

```{r fig1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Rates for Regret, Social norms and Negative affect"}
data1 <- read.csv("/cloud/project/inputs/data/Ex1.csv", header = TRUE, stringsAsFactors = FALSE, fileEncoding = "UTF-8-BOM")
str(data)

# Figure 1 replication

library(ggplot2)
library(reshape2)
library(ggplot2)
library(readr)
library(dplyr)

data <- data.frame(
  Category = c('Regret', 'Social norms (injunctive)', 'Social norms (descriptive)', 'Negative affect'),
  Smith_Routine = c(7.9, 4.7, 90.6, 7.3),
  Jones_Exception = c(92.1, 95.3, 9.4, 92.7)
)

long_data <- melt(data, id.vars = 'Category', variable.name = 'Group', value.name = 'Percentage')

ggplot(long_data, aes(x = Category, y = Percentage, fill = Group)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  scale_fill_grey(start = 0.8, end = 0.2) + # Adjust the grey scale to match the example
  theme_minimal() +
  labs(
    fill = "", 
    x = "",    
    y = ""     
  ) +
  theme(
    legend.position = "bottom", 
    axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis labels if necessary
    plot.title = element_text(hjust = 0.5) # Center the plot title if you add one
  )
print(ggplot)
```

## Demographic Factors

Figure \@ref(fig:fig2) is a replication of Figure 2 in the original paper, and shows the Proportions for perceived regret and luck.This graph compares the relationship between regret and luck, with 81% of 277 respondents believing that Mr. White, who had just changed to a new route home and experienced an accident, would feel more regretful.
```{r table1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Rates for Regret, Social norms and Negative affect"}

table2 <- read.csv("/cloud/project/inputs/data/table2.csv", header = TRUE, stringsAsFactors = FALSE, fileEncoding = "UTF-8-BOM")
print(table2)
```
 As for luck, Mr. White (exception) was considered less lucky than Mr. Adams (routine), with 114 participants (33%) viewing Mr. Adams as less lucky, whereas 228 (66%) perceived Mr. White to be less lucky.

```{r fig2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Trends in Birth Rates by Population Subgroup"}

library(ggplot2)
library(reshape2)

df <- read.csv("/cloud/project/inputs/data/table2.csv", header = TRUE, stringsAsFactors = FALSE)

# Clean up the data by removing unnecessary 'Counts' columns and renaming the remaining
df_clean <- df[, -c(2, 4)]
names(df_clean) <- c("Category", "Regret", "Luck")

# Melt the data frame to have 'Category', 'Variable', and 'Value' columns
df_melted <- melt(df_clean, id.vars = "Category", variable.name = "Outcome", value.name = "Percentage")

# Plot with ggplot2
p <- ggplot(df_melted, aes(x = Outcome, y = Percentage, fill = Category)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.7)) +
  scale_fill_manual(values=c("Routine Adams" = "black", "Exception White" = "grey")) +
  labs(x = "", y = "Percentage", fill = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")

# Print the plot
print(p)
```

```{r proportions, echo=FALSE, warning=FALSE, message=FALSE}
this.dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
setwd(this.dir)

# load our dataset 
data <- read.csv("/cloud/project/inputs/data/osf-past-normality-regret-replication-exp1-data.csv", header = TRUE, stringsAsFactors = FALSE, fileEncoding = "UTF-8-BOM")
str(data)

data$Sc1_regret<- factor(data$Sc1_regret)
data$sc1_socnorms1<- factor(data$sc1_socnorms1)
data$sc1_socnorms2<- factor(data$sc1_socnorms2)
data$sc1_combinednorms<- factor(data$sc1_combinednorms)

# Let's label the values better, so it's easier to understand the output.
data$Sc1_regret <- ordered(data$Sc1_regret, levels = c(1,2), labels = c("Exception Jones", "Routine Smith"))
data$sc1_socnorms1 <- ordered(data$sc1_socnorms1, levels = c(1,2), labels = c("Exception Jones", "Routine Smith"))
data$sc1_socnorms2 <- ordered(data$sc1_socnorms2, levels = c(1,2), labels = c("Exception Jones", "Routine Smith"))
data$sc1_combinednorms <- ordered(data$sc1_combinednorms, levels = c(1,2), labels = c("Exception Jones", "Routine Smith"))

data$Sc3condition1 <- 0
data$compensationagg1 <- 0
data$regretagg1 <- 0
for (i in 1:nrow(data)){
  if (!is.na(data$Sc3_C1_text[i])){
    data$Sc3condition1[i] <- 1
    data$compensationagg1[i] <- data$sc3_c1_compensation[i]
    data$regretagg1[i] <- data$sc3_c1_regret[i]
  }
  else if (!is.na(data$Sc3_C2_text[i])){
    data$Sc3condition1[i] <- 2
    data$compensationagg1[i] <- data$sc3_c2_compensation[i]
    data$regretagg1[i] <- data$sc3_c2_regret[i]
  }
  else if (!is.na(data$Sc3_C3_text[i])){
    data$Sc3condition1[i] <- 2
    data$compensationagg1[i] <- data$sc3_c3_compensation[i]
    data$regretagg1[i] <- data$sc3_c3_regret[i]
  }
  else {
    data$Sc3condition1[i] <- NA
    data$compensationagg1[i] <- NA
    data$regretagg1[i] <- NA
  }
}

#value labels
data$Sc3conditionl1<-factor(data$Sc3condition1,levels = c(1,2), labels=c("Routine", "Exception"))

# let's have a look at this
table(data$Sc3conditionl1)

#Adjust Copensation-Scale to original paper (Values from 0 to 10, instead of 1 to 11)
data$compensationaggrecoded1 = data$compensationagg1-1

#Adjust Regret Scale (Values 1 to 5, indead of 0 to 4)
data$regretaggrecoded1 = data$regretagg1+1

# Label Variables
names (data$compensationaggrecoded1) <- c("0", "100,000", "200,000", "300,000", "400,000", "500,000", "600,000", "700,000", "800,000", "900,000", "1,000,000")
names (data$regretaggrecoded1) <- c("no regret", "weak regret", "medium regret", "strong regret", "very strong regret")

#######
# Compensation
########

# what was the effect in the original study?
ci.smd(ncp=2.17, n.1=58, n.2 = 105, conf.level=0.95)

exceptioncombinedcompensationplot <- ggplot(data, aes(x=Sc3conditionl1, y=compensationaggrecoded1)) + 
  geom_violin() + geom_violin(trim=FALSE) + 
  scale_x_discrete(limits=c("Routine", "Exception")) +
  geom_jitter(shape=16, position=position_jitter(height=0.1, width=0.2)) +
  scale_color_grey() + scale_fill_grey() + 
  theme_apa() +
  theme(axis.title.x=element_blank(), 
        axis.text = element_text(size = 16),
        axis.title.y = element_text(size = 20),
        panel.grid.major = element_line(colour="grey", size = (0.1)),
        panel.grid.minor = element_line(size = (0.1), colour="grey"),
        text=element_text(family="Arial"))+
  xlab("Condition") +
  ylab("Compensation") +
  scale_y_continuous(minor_breaks = seq(-10, 15, 5)) +
  stat_summary(fun.data = mean_se, geom = "errorbar",
               position = position_dodge(width = 0.90), width = 0.2) +
  stat_summary(fun.y=mean, geom="point", shape=18,
               size=3)
exceptioncombinedcompensationplot

########
# Regret
########

exception_combined_regret_plot <- ggplot(data, aes(x = Sc3conditionl1, y = regretaggrecoded1)) +
  geom_violin(trim = FALSE) + # Duplicate geom_violin removed
  geom_jitter(shape = 16, position = position_jitter(height = 0.1, width = 0.2)) +
  scale_x_discrete(limits = c("Routine", "Exception")) +
  scale_color_grey() + 
  scale_fill_grey() + 
  theme_apa() +
  theme(
    axis.title.x = element_blank(), 
    axis.text = element_text(size = 16),
    axis.title.y = element_text(size = 20),
    panel.grid.major = element_line(colour = "grey", size = 0.1),
    panel.grid.minor = element_line(size = 0.1, colour = "grey"),
    text = element_text(family = "Arial")
  ) +
  xlab("Condition") +
  ylab("Regret") +
  stat_summary(fun.data = mean_se, geom = "errorbar",
               position = position_dodge(width = 0.9), width = 0.2) +
  stat_summary(fun.y = mean, geom = "point", shape = 18,
               size = 3)

# Print the plot
exception_combined_regret_plot

```
In the first graph, we can observe two distributions corresponding to the compensation under routine behavior conditions and exception conditions. The compensation amounts are represented on a scale from 0 to 10, which might indicate a range from $0 to $1,000,000. Under the routine behavior condition, the data distribution appears to be more concentrated, suggesting that most participants' compensation amounts were relatively close to each other, and the median seems to be near the middle value of the scale. Under the exception condition, the distribution of compensation data is broader, indicating a greater variance in the participants' assessments of the compensation amount.

The second graph illustrating the distribution of reported regret under routine and exception conditions.In the routine condition, the distribution of regret is narrower and more concentrated around the lower end of the scale, suggesting that participants generally felt less regret for the routine situation. The median regret rating, indicated by the horizontal line inside the violin plot, appears to be below the midpoint of the scale.In the exception condition, the regret distribution is wider at the top, indicating that there is a greater variance in how much regret participants felt, with more reports of higher levels of regret. The median regret rating in this condition is higher than in the routine condition, as seen by its position above the midpoint of the scale.

# Discussion

In Figure \@ref(fig:fig1), we saw that the birth rate enters a persistent decline after the Great Recession in 2007. However, one is left to wonder if this recession is really the event to demarcate this trend in birth rates, since it does not mirror the pattern of recessions prior as we have yet to see any increase in birth rates after the decline.

## Why is the birth rate declining? - Demographic insights

To investigate this idea more, we first turn to the trends among the various demographic subgroups. For the different age groups, the decreasing trend among younger age groups and increasing trend among older age groups points to the idea that women are choosing to have children later in life, an attitudinal change. Only the age 20-24 group seems to show a decrease at the 2007 mark that differs significantly from the years prior. One explanation is that it is the age when most people attend university, and more women are applying for tertiary education and pursuing education before having children. In fact, the National Center for Education Statistics (NCES) data shows a 17.6% increase in undergraduate enrollment among females from 2007 to 2011 [@nces], which coincides perfectly with the 19.1% decrease in birth rates in the 4 years after the Great Recession began in 2007.

For race and ethnicity, the birth rates for Black and White women remains fairly constant and only the Hispanic women show a significant decline in the years 2007-2011 before it plateaus to follow a similar trend to the other two races. @og offer assimilation as a possible explanation with evidence to support this notion from @tavernise and @Parrado2008.

For marital status, married and unmarried women have fairly unchanged birth rates around 2007, so it does not appear to be affected too much by the recession or other events around that time. @og suggests that the small decrease in birth rates for unmarried women is possibly due to the decline in teenage births (ages 15-19) that we saw in Figure \@ref(fig:demgraphs) plot A, since women are marrying later and therefore teenage women make up a larger proportion of the unmarried women population in their childbearing years [@marriage].

For education level, the groups do not appear to demonstrate any change in birth rates with the presence of the Great Recession, except for the no highschool degree group. Its rise and fall in birth rates may be related to the boom in the residential housing market leading up to the recession, where many subprime mortgages were given out, that is, loans to persons with low credit ratings, of which women with a lack of education i.e. no highschool degree would fall under [@insider]. Thus, more women in this group may have chosen to buy homes and start families, and then the recession hit and reversed this effect.

Thus across these demographic groups that we have examined, we note two things: 1) the the declining trend in the total female population of childbearing age is not paralleled equally across its subgroups of age, race and ethnicity, marital status and education level, but 2) specifically we note that younger age groups, Hispanic women and women with no highschool degree showed a significant decrease in birth rates of which only the no highschool degree mamas seem to be related to the Great Recession. We did mention earlier however, that there seems to be an attitudinal change among the women close to the recession - women having children later in life, getting married later, and more women attending colleges and universities - which leads into our discussion of cohort effects.

## Cohort effects - attitudes of a new generation

From Figure \@ref(fig:cohorttab), it was shown that the birth cohorts in years 1968-1982 are quite similar to each other whereas the cohorts born in years 1983-1997 seemed to pull more away with lower birth rates compared to its older counterparts. In fact, these two groupings of years define two distinct generations respectively: Generation X and Generation Y, or the "Millenials" [@gendef]. @millenial conducted a study comparing various lifestyle factors such as education, income, family and housing of Millenials to its prior generations and there are some stark differences.

Notably, women are delaying motherhood as according to a Pew Research Center analysis of population data, 2016 saw 48% of Millennial women having given birth to at least one child at ages 20 to 35 compared to the previous generation where 57% of women were already mothers in 2000 for the same age range [@milmom]. The marriage status for millenials also shows similar figures with 46% of Millennial women ages 25 to 37 being married compared to 57% of Generation X-ers for the same age range, and less millenial women are married for lower levels of education [@millenial]. Additionally, the generations are becoming more racially and ethnically diverse [@postmil] with the population of Hispanics less likely to be foreign-born [@og]. These support our arguments for the change in birth rate trends for women of the different age groups, marital status and race and ethnicity that was discussed in the previous section.

So it now begs the question, why are the attitudes between these two generations so different? We briefly touch on some historical events that surely would have influenced change in Generation Y.

## The Digital Revolution and Third Wave Feminism

The birth years of the millenial generation were characterized by several milestones in the Digital Revolution: the first mobile phone (1984), digital camera (1988) and the World Wide Web (1989) [@stfc]. When the first set of millenials would turn about 10 years old, CERN’s web browser software was released for public use (1991) and commercial dial up internet became available (1992) [@stfc]. In their teenage years to follow, smartphones, social media and high speed internet would be developed creating a tech-savvy generation like never before.

It also happens that Generation X coincides with the years of the second wave of feminism (1963 - 1980s) whereas the formative years of millenials coincide the years of the third wave of feminism [@vox]. The third wave of the feminist movement began with the Anita Hill sexual harassment case in 1991, followed by the "Year of the Woman" in 1992 where a record-breaking number of women candidates were elected to the House of Representatives and Senate (@cawp, as cited in @Dolan1998, p. 272). The third wave of feminism attempts to be distinctive from its predecessor, with some claims in popular literature of it to have a culture of 'girl power', be more open-minded, more racially diverse and inclusive than the second wave, although it is disputed whether some of these claims are exaggerated [@Snyder2008].

Therefore, the influence of technology and mass media that comes with it as well as the different ideologies behind feminism and women empowerment may have shaped the different attitudes we see between the more recent birth cohorts and older cohorts, and these changes may manifest in the decline we see in the birth rates.

## The decline of birth rates - not an effect of the Great Recession?

Back to Figure 1, we noted that the first time the birth rate dipped below minimum from previous years was in 2011, which, upon reflection after our discussion, is almost exactly 20 years after the commercial dial up internet became available and the third wave feminist movement began. 2012 is the year when the birth cohort 1988-1992 was ages 20-24, which is the first cohort to pull away significantly in birth rates to the previous ones. So are these other historical events indirectly responsible for the persistent decline in birth rates by influencing women's attitudes, and not the Great Recession? This is left as food for thought for now, more research would be needed to investigate this question.

# Conclusion

We explored the trend in birth rates in the US for the period 1980-2020 among the total child-bearing population, and demographic subgroups by age, race and ethnicity, marital status and level of education. Overall there is a decline, but the trend oftentimes remains constant for different subgroups and only declines significantly for specific groups such as younger age groups, Hispanic women and women with no highschool degree after 2007. The original paper by @og uses the Great Recession as the event to mark the beginning of the birth rate decline, but although we acknowledge some effect of the recession on decreasing birth rates, we conjecture that the lasting and continuous decline is greater explained by the shifing attitudes towards when and if to have children in the new generation, and that the decline would have been evident even if the recession had not occurred. However, if women are simply choosing to have children later and not to not have children at all, we should expect a pick back up in the birth rates in the coming decade as this group ages.

## Weaknesses and next steps

A weakness is that we only described a subset of the demographic variables that were presented in @og's paper and so the analysis could have been more complete. I do not claim to be an expert on the technlogical revolution or feminism, and we only provided these themes as potential explanatory factors for the change in attitude that is causing the decline in birth rates. More research into how these variables correlate with the different generational groups or birth cohorts is needed to make stronger statistical claims, such as by collecting data on women's beliefs and possibly constructing regression models if appropriate.

\newpage

# References